\section{BDI Agent}
The agent operates on a continuous cycle driven by the \texttt{sensing} socket event emitted by the server.
Rather than relying on a static script, the agent dynamically reacts to the environment through a strict, four-stage pipeline.
Upon receiving a \texttt{sensing} event, the following sequence executes:
\begin{enumerate}
    \item \textbf{Perception (\texttt{perceive()}):} The agent integrates the new sensory data into its belief state trough a Belief Revision process (detailed in Section~\ref{sec:belief}).
    The actions we perform that modify the belief state automatically updates the beliefs, without having to wait for the next sensing event.

    \item \textbf{Deliberation (\texttt{deliberate()}):} Using its updated beliefs, the agent generates potential goals (desires) and calculates a score for each.
    It then rebuilds its intention queue, prioritizing the most valuable task (detailed in Section~\ref{sec:desires}).

    \item \textbf{Planning (\texttt{Planner.plan()}):} The planner extracts the highest-ranked intention from the queue and translates it into a concrete sequence of
    executable, typed steps (detailed in Section~\ref{sec:planning}).

    \item \textbf{Execution (\texttt{Executor}):} Finally, the planned actions are asynchronously emitted to the server via socket events.
    To maintain responsiveness, the executor triggers an action at each server clock tick, instead of waiting for a perception update or an acknowledgment from the server.
\end{enumerate}

\subsection{Belief Revision}
\label{sec:belief}
The agent's belief state relies on two foundational utility data structures to manage temporal observations,
which in turn feed into three specialized sub-systems: \texttt{AgentBeliefs}, \texttt{MapBeliefs}, \texttt{ParcelBeliefs}.

\paragraph{Tracker} \texttt{Tracker} stores a single latest observation $\langle v, t_{seen} \rangle$ per entity \texttt{e}.
On each sensing report, \texttt{update(e, v)} overwrites the existing entry.
The confidence of a tracked observation decays exponentially over time:
\begin{equation}
c(t) = \exp \left(-\frac{t - t_{seen}}{\tau}\right)
\label{eq:confidence}
\end{equation}
where $\tau$ is the exponential time constant (the duration after which confidence drops significantly).

When the current sensing window covers an entity's last-known position, but the entity is absent from the report,
\texttt{invalidateAtSensedPositions} nullifies that position without removing the record entirely.

\paragraph{Memory}
While the tracker maintains only the latest state, \texttt{Memory} is a supplementary store that maintains a bounded temporal history of observations per entity, automatically evicting entries older than a defined time threshold.

\subsubsection{Agent beliefs}
\texttt{AgentBeliefs} stores the latest state of every observed agent via \texttt{Tracker}, and a bounded position history for each enemy via \texttt{Memory}.
It also maintains a spatial heat field over the map, aggregating enemy presence across all tracked observations.

\subsubsection{Map beliefs}
\texttt{MapBeliefs} stores the static tile layout, a set of transient blocked tiles each with a configurable TTL, and soft traversal penalties on individual tiles.

Upon map receipt, it identifies and permanently seals \emph{sink tiles}: tiles that an agent can enter but from which no valid pickup-and-deliver loop exists.
This situation arises on maps where directional conveyor tiles create one-way corridors: an agent may be carried into a region
from which it can never return to a spawn tile, or from which no delivery tile is reachable.

To detect these dead zones, the map is modelled as a directed walkability graph where conveyors enforce one-way edges.
Tarjan's algorithm is then applied to find all Strongly Connected Components (SCCs).
An SCC is considered \emph{safe} only if it contains at least one spawn tile \emph{and} at least one delivery tile (i.e.\ it supports the full pickup-and-deliver cycle).
All walkable tiles outside a safe SCC are reclassified as walls, preventing the planner from ever routing the agent into an unrecoverable region.


\subsubsection{Parcel beliefs}
This component tracks the availability and value of delivery targets.
Parcels within the current sensing window receive their authoritative reward from the server.
Once out of sight, parcel rewards decay linearly at $-1$ per server-defined interval $\Delta_t$.
Each parcel maintains an independent decay clock to prevent drift accumulation.
Given a parcel last observed at time $t_0$ with reward $r_0$, its estimated reward at time $t$ is:
\begin{equation}
r(t) = r_0 - \left\lfloor\frac{t - t_0}{\Delta_t}\right\rfloor.
\label{eq:decay}
\end{equation}
When $r(t) \leq 0$, the parcel is considered expired and is deleted from the agent's beliefs.

\subsubsection{Crate beliefs}
\texttt{CrateBeliefs} uses the same \texttt{Tracker} utility described above to store the latest known position for each crate.
On every sensing update, observed crate positions are registered and \texttt{invalidateAtSensedPositions} nullifies any entry whose tile falls within
the current sensing window but is absent from the report, reflecting that the crate has been pushed.

\subsection{Desires and Intention Selection}
\label{sec:desires}
For the BDI layer, we define a fixed set of 3 core desire types, each parameterized by a target tile:
\begin{itemize}
    \item \textbf{ReachParcel}: navigate to and collect a visible parcel;
    \item \textbf{DeliverParcel}: navigate to a delivery tile while carrying parcels;
    \item \textbf{Explore}: navigate to an unobserved spawn tile when idle (i.e., no reachable parcels or deliveries).
\end{itemize}

\subsubsection{Queue reconstruction}
At each deliberation step, the intention queue is fully reconstructed from the current belief state.
The queue is sorted by priority first, then by descending score within each desire type.
Both \texttt{ReachParcel} and \texttt{DeliverParcel} desires have a priority of 1, while \texttt{Explore} has a lower priority of 0.
This ensures that the agent always prefers tangible parcel-related goals over exploration when any are available.

\subsubsection{Dynamic Scoring}
The agent computes a dynamic score for each desire to guide intention selection.
The score is a real-valued estimate of the desire's expected utility, incorporating factors such as the parcel's reward, the distance to the target, and the presence of competitors.

\paragraph{Distance Penalty}
Each desire's score is inversely proportional to the estimated distance to its target, calculated using the agent's current beliefs.
The agent computes the shortest path distance $d$ from its current position to all other tiles using a Breadth-First Search (BFS).
We will refer to this distance as $d$ throughout the rest of the section. (Note: when $d = 0$, we assign an infinite score).

\paragraph{ReachParcel}
Let $r_p$ denote the current estimated reward of a target parcel $p$, and $C$ the set of parcels the agent is currently carrying.
The score $S$ for a \texttt{ReachParcel} desire targeting parcel $p$ is calculated as:
\begin{equation*}
  S = c_p \cdot \frac{r_p + \sum_{c \in C} r_c}{d} \cdot \gamma
\end{equation*}
where $c_p$ is the confidence in the target parcel's last known position (Eq.~\ref{eq:confidence}, with $\tau = \tau_c$), reflecting how stale the parcel observation is.
A parcel that was just sensed has $c_p \approx 1$ and is scored at full value; one that has been out of sight for several half-lives is discounted accordingly.

$\gamma$ is a race factor that de-prioritizes parcels likely to be collected by competitors. It is computed as:
\[ \gamma = \min\left(1, \frac{d_{\text{enemy}}}{d}\right) \]
where $d_{\text{enemy}}$ is the distance from the nearest tracked enemy to the parcel.

Incorporating the total reward of currently carried parcels ($\sum_{c \in C} r_c$) into the score allows the agent to evaluate the utility of
picking up a new parcel against the potential value of delivering existing ones. This effectively balances the agent's acquisition and delivery goals.

\paragraph{DeliverParcel}
For a \texttt{DeliverParcel} desire, the objective is to secure the rewards of the parcels currently held by the agent. The score $S$ is computed as the total estimated reward of all carried parcels $C$, inversely scaled by the shortest path distance $d$ to the delivery tile:
\begin{equation*}
  S = \frac{\sum_{c \in C} r_c}{d}
\end{equation*}

\paragraph{Explore}
The \texttt{Explore} desire serves as a fallback to locate newly spawned parcels. We first define the normalized observation age $\tau$:
\begin{equation*}
  \tau = \min\left(1, \frac{\Delta t}{T_{\text{spawn}}}\right)
\end{equation*}
where $\Delta t$ is the time since the target spawn tile $x$ was last observed, and $T_{\text{spawn}}$ is the environment's spawn interval.

The cluster weight $w$ quantifies the structural density of the target's surrounding spawn area. It is computed as the sum of inverse distances to all neighbor spawn tiles within the observation range $R_{\text{obs}}$:
\begin{equation*}
  w = \sum_{y \in OST} \frac{1}{\delta(x, y) + 1}
\end{equation*}
where $OST$ is the set of spawn tiles such that their Manhattan distance $\delta(x, y) \le R_{\text{obs}}$.

The enemy heat $h$ actively steers the agent away from sectors currently crowded by competitors.
It is accumulated from the history of tracked enemy observations $E$, combining spatial and temporal decay, and weighted by each enemy's track confidence $c_e$ (Eq.~\ref{eq:confidence}):
\begin{equation*}
  h = \sum_{e \in E} c_e \exp\left(-\frac{d_e^2}{2\sigma^2}\right) \exp\left(-\frac{\Delta t_e}{\tau_h}\right)
\end{equation*}
where $d_e$ is the spatial distance to the past enemy observation, $\Delta t_e$ is the observation's age, $\sigma$ controls the spatial spread, $\tau_h$ is the temporal decay rate, and $c_e = \exp(-(t - t_{seen,e})/\tau_c)$ is the confidence in the enemy's last known position.

Finally, the score for an \texttt{Explore} desire is computed as:
\begin{equation*}
  S = \frac{w \cdot \tau}{d \cdot (1 + h)}
\end{equation*}

\subsection{Planning: A* Navigation and Collision Handling}
\label{sec:planning}

For pathfinder, the agent employs a standard A* search algorithm with a Manhattan distance heuristic, which is admissible in our grid-based environment.
The planner incorporates dynamic walkability checks based on the agent's current beliefs, treating temporarily blocked tiles as impassable.
Statically, walls and out-of-bounds tiles are always rejected; conveyor tiles are additionally direction-sensitive, blocking movement against their flow.

Plans generated by the A* algorithm are defined as sequences of typed actions: directional \texttt{move} steps followed by a terminal \texttt{pickup} or \texttt{putdown} step.
An existing plan is reused across deliberation steps as long as the active desire type and target remain unchanged, and all remaining steps successfully pass walkability re-validation.

\paragraph{Collision Handling}
Prior to executing a move step, the agent verifies the availability of the target tile. A tile is considered occupied, and thus temporarily blocked, under two conditions:
either an enemy is currently observed on it, or an enemy's predicted next position matches it with sufficient confidence.

When the next planned tile is occupied, the collision manager applies a three-stage escalation:
\begin{enumerate}[noitemsep]
  \item \textbf{Detour}: replan with the blocked tile treated as a wall.
  This is accepted only if the detour adds at most a predefined threshold of $\Delta_\text{th}$ steps over the remaining plan.
  \item \textbf{Wait}: if no acceptable detour exists, pause execution for a random duration drawn from a predefined uniform distribution.
  \item \textbf{Block and replan}: after the wait expires, or after a specified limit $L$ of repeated invalidations,
  mark the tile as temporarily blocked for a randomized time-to-live (TTL) duration and completely replan toward the same goal.
\end{enumerate}

This escalation applies only to enemies; teammate collisions are resolved deterministically by agent ID, with the lower ID conceding.