\section{LLM-Based Agent Layer}
\label{sec:llm}
The LLM layer functions as a thin wrapper around the underlying BDI agent.
It interprets high-level textual commands from a mission controller, utilizing tools exposed by the BDI core to inject new desires,
adjust scoring rules, or initiate team coordination.

The mission controller can issue requests that fall into three broad categories: direct navigation goals, persistent behavioural adjustments, and team-level control.
Each category is handled through a different mechanism in the BDI core, chosen to integrate naturally with the existing deliberation loop rather than bypassing it.

\subsection{Adapting Behavior Through Goal Injection}

Navigation requests, more specifically, moving to a tile (\emph{go-to}) or delivering at a specific location (\emph{deliver-at}), are translated into \texttt{REACH\_TILE} and \texttt{DELIVER\_PARCEL} desires, respectively. 
These injected goals enter the BDI intention queue and persist for a specified Time-To-Live (TTL), allowing them to be continuously re-evaluated against existing tasks until completion or expiry.

Both injected desires utilize the standard BDI rate formulas (Section~\ref{sec:desires}), 
but incorporate configurable rewards from the mission controller to dictate their competitive priority:

\begin{itemize}
    \item \textbf{\texttt{REACH\_TILE}:} Scored identically to \texttt{ReachParcel}:
    \begin{equation*}
      S = \frac{r_{\text{inj}} + \sum_{c \in C} r_c}{d}
    \end{equation*}
    where $r_{\text{inj}}$ is the injected reward. Setting this reward equal to a typical parcel value allows the navigation goal to compete on equal footing with standard parcel collection.

    \item \textbf{\texttt{DELIVER\_PARCEL}:} Similarly, applies the standard rule-adjusted scoring for deliveries (Section~\ref{sec:scoring_rules}), including an optional additive bonus after the rule combination to successfully bias the agent toward the requested destination tile.
\end{itemize}

\subsection{Adapting Strategy via Rule Injection}
\label{sec:scoring_rules}
The \texttt{RuleStore} maintains a persistent registry of LLM-defined scoring rules.
Each rule encodes a state condition alongside a linear effect: a multiplier $m$ and an additive offset $a$.

During each deliberation cycle, any matching rules are applied to the baseline scores from Section~\ref{sec:desires} as a linear combination.
This scales or shifts the priorities without overwriting the foundational scoring logic.

These rules evaluate three primary state variables: the individual parcel reward, the current carry count, and the target delivery tile.
Their effects compose directly onto the reward aggregates utilized in the desire formulas.

\paragraph{ReachParcel:}
The baseline score (Section~\ref{sec:desires}) is linearly adjusted according to the \emph{estimated stack count} post-pickup: the hypothetical carry count $|C|+1$.
This allows a rule to incentivize picking up a specific parcel only if it achieves a desired total stack size, leaving all other pickup valuations unchanged.

\paragraph{DeliverParcel:}
The delivery score is linearly modified based on the current carry count $|C|$ and, optionally, the specific target tile.
A stack rule keyed to $|C|$ dynamically alters the attractiveness of delivery depending on the agent's current payload, while a tile-specific rule can shepherd the agent toward a preferred drop-off location.
An independent bonus can also be applied directly via the \texttt{request\_putdown\_at} tool, biasing delivery toward a specific tile regardless of any active rules.

\paragraph{Explore:}
In the absence of an active stack rule, \texttt{Explore} maintains its default priority of 0 and functions as outlined in Section~\ref{sec:desires}.

However, when a positive stack rule is triggered, \texttt{Explore} is elevated to priority 1, forcing it to compete directly against active parcel and delivery goals.
Its score is subsequently scaled by the anticipated reward of completing the stack, incentivizing the agent to actively search for missing parcels rather than settling for a premature delivery.
This behavior emerges naturally from the linear weighting system: the LLM is not required to issue an explicit exploration command, as simply registering a stack rule is sufficient to trigger the shift.

\paragraph{Traversal Penalties:}
Traversal penalties operate at the planning layer rather than the deliberation layer.
Unlike scoring rules, which modulate desire priority during deliberation, a traversal penalty injects an additive cost onto the A* edge leading onto a specific tile.
When the planner evaluates candidate paths, it adds this cost to any step landing on a penalized tile, causing the agent to prefer a detour whenever the detour length is shorter than the penalty,
without completely forbidding the option of traversing the tile if it becomes necessary.

\subsection{Coordination}
\label{sec:coordination}

Coordination in our system is fully decentralized and message-driven: there is no central authority, and every joint decision emerges from peer-to-peer exchanges.
Since only the LLM-driven agent is wired to the mission controller, it is the one that interprets admin queries and is therefore the sole agent able to issue proposals.

\paragraph{Beacon Communication:}
To handle the critical need for up-to-date position information in coordination maneuvers, each agent repeatedly broadcasts its current position 
to all configured teammates as a \texttt{position\_beacon} message at a fixed interval. 
This continuous broadcast is essential for enabling reliable coordination, even with teammates that fall outside of an agent's direct sensory range.

\paragraph{Peer Communication:}
Simple agent-level commands (navigation goals, scoring rules, traversal penalties) follow a uniform two-step pattern: 
the injection is applied to the issuing agent locally, then packaged as a peer-injection message and broadcast to all teammates.

Commands that carry stronger team-synchronisation requirements (rendezvous and traffic control) instead use a two-phase commit protocol described below.

\paragraph{Two-Phase Commit Protocol:}
Commands that carry stronger team-synchronization requirements utilize a generalized two-phase commit (2PC) protocol. 
This ensures the team only commits to a synchronized interruption of their current tasks when it is collectively optimal to do so.

\begin{itemize}
    \item \textbf{Phase 1: Propose.} The proposing agent broadcasts a \emph{propose} message carrying a unique round identifier (\texttt{rid}) and the command parameters. 
    Each teammate evaluates the proposal based on its current local state and replies with a \emph{vote} message indicating its accept or reject decision for that \texttt{rid}.
    \item \textbf{Phase 2: Commit.} After a fixed collection window, the proposer checks for unanimity. 
    If every expected teammate voted \emph{accept}, the proposer broadcasts a \emph{commit} message, and all agents (including the proposer) execute the command. 
    If any vote is missing or negative, the proposer broadcasts \emph{abort}, and the command is discarded globally.
\end{itemize}

\paragraph{Protocol Evaluation and Execution:}
We specifically apply this 2PC protocol to coordination maneuvers that force an agent to hold its position, namely \texttt{request\_rendezvous} and \texttt{request\_red\_light}. 
In both cases, the successfully committed action results in the injection of a \texttt{HOLD\_TILE} desire.

During Phase 1, the proposal payload offers a set of candidate hold tiles: for rendezvous, the tiles around the meeting point; for traffic control, the tiles lying on alternate rows or columns. 
A teammate evaluates this proposal by scoring the candidates and keeping the best \texttt{HOLD\_TILE} score, reusing the reward-over-distance formulation from Section~\ref{sec:desires}. 
To keep the comparison fair, the carried value is first discounted by the parcel-reward decay expected during the wait—either until the slowest teammate arrives for a rendezvous, or over the full duration for traffic control. 
The teammate votes \emph{accept} only if this discounted hold score outweighs its current active desires.

Once committed, holds are released based on their specific command semantics. 
A rendezvous hold releases itself automatically once position beacons confirm all teammates have gathered within the agreed radius, requiring no further messages. 
Conversely, a traffic-control hold persists until its time limit expires or the mission controller issues an explicit green-light command that releases the whole team at once.

\subsection{Cooperation}

Beyond reactive primitives, the team periodically evaluates a higher-level cooperation strategy.
Every $k$ seconds, the LLM agent initiates a cooperation round by requesting teammates' beliefs.
From the replies it assembles a compact context for an LLM query, drawing on three sources: a \emph{roster} listing each agent's position, carried load, and any sensed enemies; 
a deterministic \emph{geometry} summary of the spawn clusters and delivery zones; 
and a \emph{performance} record holding the average score each strategy has earned per round over a recent sliding window.
The prompt is steered with a handful of \emph{few-shot examples}, each pairing a representative scenario with its expected decision.
It is further guided by \emph{chain-of-thought} prompting: the model is asked to reason step by step and to return a structured response that states its rationale before committing to a strategy.

\paragraph{Strategies:}
Two cooperative strategies are available alongside a default non-cooperative baseline.
In \texttt{ZONAL\_RELAY}, one agent clears the spawn region and drops parcels at a fixed midpoint tile, while the other ferries them to delivery tiles.
In \texttt{OPPORTUNISTIC}, agents forage independently and hand off parcels dynamically whenever a pickup occurs.

\paragraph{Roles:}
Each strategy assigns agents specific roles executed via a deterministic finite-state machine (FSM).
This FSM supersedes the standard BDI deliberation, emitting a restricted set of state-gated desires to enforce the intended cooperative behavior.
The spatial targets for these roles are computed deterministically from map geometry; the exchange tile is the reachable cell closest to the midpoint between spawn and delivery clusters, and each agent is assigned a distinct adjacent approach cell.
This position is injected as a high-reward \texttt{REACH\_TILE} desire until the FSM transitions.

\paragraph{Handoff:}
Both strategies culminate in a handoff.
In \texttt{ZONAL\_RELAY}, roles are fixed and the exchange loops indefinitely.
In \texttt{OPPORTUNISTIC}, roles are fluid: the agent that acquires a parcel proposes a handoff via a two-phase commit protocol, becoming the \emph{passer} once the partner commits as the \emph{receiver}.
The partner accepts only if the meeting tile is reachable.
When the mission controller issues an explicit directive, the LLM does not alter this core machinery; it simply applies a reward bonus to the handoff desire to make the exchange more highly prioritized.