\section{PDDL-Based Crate Navigation}
\label{sec:pddl}
Some maps in the Deliveroo.js environment contain movable crates that can obstruct navigable paths. 
An agent may push a crate one tile in any cardinal direction, subject to the constraints outlined in Section~\ref{sec:planning}. 
We model this crate manipulation problem using PDDL (Planning Domain Definition Language), invoking it as a fallback mechanism when the standard A* planner fails to find an unobstructed route.

\subsection{Domain Overview}
The \texttt{deliveroo-crates} domain is formulated in STRIPS and relies on two primary object types: \texttt{tile} and \texttt{crate}. The complete state required for crate-aware navigation is captured by the following predicate set:
\begin{itemize}[noitemsep]
  \item \texttt{(at ?t)}: The agent is located on tile $t$.
  \item \texttt{(adj-up/down/left/right ?from ?to)}: Defines the valid directed spatial relationships between adjacent tiles for movement or pushing.
  \item \texttt{(crate-at ?c ?t)}: Crate $c$ occupies tile $t$.
  \item \texttt{(crate-free ?t)}: Tile $t$ contains no crate, making it available for entry or as a push destination.
  \item \texttt{(crate-space ?t)}: Tile $t$ can structurally support a crate (e.g., it is not a wall or a conveyor belt).
\end{itemize}

The domain defines eight parameterized actions: four directional \emph{move} actions and four directional \emph{push} actions.
\begin{itemize}
    \item A move action allows the agent to step into an adjacent \texttt{crate-free} tile. 
    \item A push action allows the agent to displace a crate one tile in the specified direction, requiring two preconditions:
    \begin{enumerate}[label=(\roman*)]
        \item the agent is positioned adjacent to the crate on the opposite side of the push direction, and
        \item the target destination is a valid \texttt{crate-space} and is currently \texttt{crate-free}. 
    \end{enumerate}
    The execution of a push action simultaneously relocates the crate to the target tile and moves the agent onto the crate's former tile, updating the \texttt{crate-free} state of the involved tiles accordingly.
\end{itemize}

\subsection{Planning With PDDL}
PDDL planning is invoked exclusively when standard A* pathfinding fails and the agent's belief system indicates that crates obstruct the route (Section~\ref{sec:planning}).
This detection occurs in two phases. First, an A* search is executed, treating all crate-occupied tiles as impassable. 
If this search fails, a second A* search is performed that entirely ignores crate positions, yielding a structurally valid but potentially obstructed path. 
Any crates from the agent's current beliefs that intersect this path are collected; their presence confirms that the route is blocked by movable obstacles rather than static map geometry.

A PDDL problem instance is subsequently constructed from the agent's current beliefs.
The \texttt{:init} state is populated with the following:
\begin{itemize}[noitemsep]
    \item \texttt{(at t\_X\_Y)}, where $(X, Y)$ is the agent's current position;
    \item adjacency facts for every pair of connected non-wall tiles, covering all four cardinal directions;
    \item \texttt{(crate-at ?c ?t)} and \texttt{(crate-free ?t)} facts for each believed crate position;
    \item \texttt{(crate-space ?t)} facts for all structurally valid positions (non-wall, non-conveyor tiles).
\end{itemize}
The \texttt{:goal} section comprises a single ground atom, \texttt{(at t\_TX\_TY)}, where $(TX, TY)$ represents the target position.
Consequently, the solver is tasked with generating a \textit{complete} sequence of \texttt{move} and \texttt{push} actions from the agent's 
current location to the final destination, clearing any obstructing crates along the way.
The resulting plan is translated back into the agent's internal action representation, stored in the plan buffer, and executed step-by-step until completion or failure.

However, this formulation suffers from a fundamental performance bottleneck: the PDDL solver is slow for the interactive demands of the environment.
An initial optimization involved transitioning to a local solver, which eliminated the overhead of remote API requests and constant problem serialization. 
A more structural optimization, caching solved plans, is unfortunately undermined by the formulation itself. 
Because the cache key must incorporate both the start and goal tiles, even a trivial change in the agent's position alters the key and triggers a cache miss, 
despite the crate-clearing maneuver remaining identical regardless of where the agent initiated the request.

\subsection{A Smart Solution: Problem Restatement \& Caching}
The root cause of this inefficiency is that PDDL is only strictly necessary for navigating the crate cluster itself; 
the initial and final segments can be efficiently handled by standard A* navigation.

To resolve this, the problem is reformulated so that the PDDL solver is only responsible for the segment of the journey that requires crate manipulation.
The overall plan is decomposed into three distinct segments:
\begin{itemize}[noitemsep]
    \item Move from the current position to the cluster entry tile (A*);
    \item Navigate through the crate cluster, from entry to exit (PDDL);
    \item Move from the cluster exit tile to the final goal (A*).
\end{itemize}
Under this paradigm, the PDDL problem is anchored to the cluster's entry tile rather than the agent's live position.
Computed maneuvers are stored in a FIFO cache, utilizing the tuple $(\text{entry},\, \text{exit},\, \{\text{crate positions}\})$ as the key.