\section{LLM Agent Layer}

% ----- LLM Layer Overview --------------------------------------
\begin{frame}{LLM Layer Overview}
  \begin{itemize}
    \item Thin wrapper around the BDI core
    \item Listens and interprets mission-controller chat messages
    \item Modifies the BDI behaviour accordingly
    \item Four command categories: \textbf{goal injection}, \textbf{rules}, \textbf{coordination} and \textbf{cooperation}
  \end{itemize}
\end{frame}

% ----- Goal Injection ------------------------------------------
\begin{frame}{Goal Injection}
  \begin{exampleblock}{Trigger Examples}
    \begin{itemize}
      \item \textit{``Move to coordinate (4,7) and you get +10pts''}
      \item \textit{``Drop a package in the leftmost tile to get 5pt''}
    \end{itemize}
  \end{exampleblock}
  
  \vspace{1em}
  
  \begin{itemize}
    \item LLM can inject new goals into the BDI agent (e.g. \texttt{go-to}, \texttt{deliver-at})
    \item Compete with existing BDI desires; may be accepted or rejected
    \item Persist and re-evaluated until executed or TTL expires
    \item \textbf{Peer-injection}: the LLM applies the command locally, then repackages it and broadcasts it
  \end{itemize}
\end{frame}

% ----- Goal Injection (diagram) --------------------------------
\begin{frame}{Goal Injection: Local + Peer Broadcast}
  \centering
  \resizebox{0.92\textwidth}{!}{\input{../../figures/goal-injection.tex}}
\end{frame}

% ----- Rule Injection ------------------------------------------
\begin{frame}{Rule Injection}
  \begin{exampleblock}{Trigger Examples}
    \begin{itemize}
      \item \textit{``Deliver stacks of exactly 3 parcels at a time to double the reward''}
      \item \textit{``Every time you deliver in (x1,y1) you get 0 pts''}
    \end{itemize}
  \end{exampleblock}
  
  \vspace{1em}
  
  \begin{itemize}
    \item \texttt{RuleStore}: Defines a condition, multiplier ($m$), and offset ($a$)
    \item \textbf{Effect}: Dynamically scales and shifts the BDI score for any desire matching the condition
    \item \textbf{Special case}: The \textbf{Stack rule} automatically elevates \textbf{Explore} to priority~1
  \end{itemize}
\end{frame}

% ----- Coordination: Position Sharing --------------------------------------
\begin{frame}{Coordination: Position Sharing}
  \begin{itemize}
    \item \textbf{Beacons}: each agent broadcasts position at fixed interval
    \item Ensures up-to-date peer positions beyond sensing radius
    \item Needed for coordination when agents are outside each other's sensing radius
  \end{itemize}
\end{frame}

% ----- Coordination: 2-Phase Commit -----------------------------------
\begin{frame}{Coordination: 2-Phase Commit}
  \begin{exampleblock}{Trigger Example}
    \begin{itemize}
      \item \textit{``Move both agents to the neighborhood of position (x,y) within a maximum distance of 3, and have them wait for each other. You will receive 500pts.''}
    \end{itemize}
  \end{exampleblock}
  
  \vspace{1em}
  
  Used for complex coordination tasks (commit together or not at all):
  \begin{itemize}
    \item \textbf{Phase 1 - Propose}: proposer broadcasts candidate hold tiles; each agent scores them
    \item \textbf{Vote}: \emph{accept} if the hold outweighs current desires; \emph{reject} if current desires are better
    \item \textbf{Phase 2 - Decide}: unanimous accept $\to$ commit (all inject); any miss or reject $\to$ abort
  \end{itemize}
\end{frame}

% ----- Coordination: 2-Phase Commit (diagram) ------------------
\begin{frame}{Coordination: 2-Phase Commit}
  \centering
  \resizebox{0.95\textwidth}{!}{\input{../../figures/two-phase-commit.tex}}
\end{frame}

% ----- Cooperation: Motivation -------------------------
\begin{frame}{Cooperation: Motivation}
  \begin{exampleblock}{Origin: Admin Directive}
    \textit{``If a parcel is initially picked up by one agent and later delivered by the other agent, you will receive a 200 points bonus.''}
  \end{exampleblock}

  \vspace{1em}
  \begin{itemize}
    \item A one-shot \emph{check-and-coordinate} is insufficient: requires dynamic pickups, meeting tile computations, and joint commitments.
    \item[$\Rightarrow$] Solution: Dedicated \textbf{roles} with specific behaviours, agreed upon via a \textbf{2-phase commit} (2PC).
    \item Generalised into a reusable framework of \textbf{strategies} and \textbf{roles}.
  \end{itemize}
\end{frame}

% ----- Cooperation: Generalisation -------------------------
\begin{frame}{Cooperation: Generalisation}
  \begin{itemize}
    \item A \textbf{cooperation round} opens every $\Delta t$ seconds (or immediately upon a mission directive).
    \item Agents request teammates' beliefs to build a \textbf{shared compact context}:
  \end{itemize}
  
  \begin{block}{Context State}
      \begin{itemize}
        \item \textbf{Roster}: Positions, payloads, sensed enemies.
        \item \textbf{Map Geometry}: Spawn/delivery clusters, separation, corridor flags.
        \item \textbf{Performance}: Average score/round per strategy (sliding window).
      \end{itemize}
  \end{block}

  \begin{itemize}
    \item[$\Rightarrow$] The \textbf{LLM} evaluates this context to select the optimal shared strategy for the team.
  \end{itemize}
\end{frame}

% ----- Cooperation: Strategies ---------------------------------
\begin{frame}{Cooperation: Strategies}
  \begin{itemize}
    \item \texttt{ZONAL\_RELAY}: One agent sweeps the spawn region and drops at a fixed midpoint; the companion ferries parcels to delivery.
    \item \texttt{OPPORTUNISTIC}: Agents forage independently and hand off parcels dynamically upon pickup.
    \item \texttt{NONE}: Fully autonomous baseline (no cooperation).
  \end{itemize}
\end{frame}

% ----- Cooperation: Roles --------------------------------------
\begin{frame}{Cooperation: Roles}
  \begin{itemize}
    \item Each strategy assigns every agent a specific \textbf{role}.
    \item Standard BDI desire generation is \emph{blocked}.
    \item A role-specific \textbf{FSM} generates desires dynamically based on the shared context and internal state.
    \item Meeting points and positioning are computed \emph{deterministically} from the map's geometry.
  \end{itemize}
\end{frame}

% ----- Cooperation: Handpass -----------------------------------
\begin{frame}{Cooperation: Handpass}
  \begin{columns}
    \column{0.60\textwidth}
      Both strategies culminate in a \textbf{handpass}:
      \vspace{0.5em}
      \begin{itemize}
        \item \texttt{ZONAL\_RELAY}:\\ The exchange loops indefinitely at the designated midpoint.
        \vspace{0.5em}
        \item \texttt{OPPORTUNISTIC}:\\ The agent acquiring a parcel proposes a handpass via 2PC.
      \end{itemize}
      
    \column{0.38\textwidth}
      %#TODO: uncomment this for video demo
      %\animategraphics[autoplay,loop,width=\linewidth]{12}{../../media/cooperation/frame-}{1}{196}
      \begin{block}{}
        \centering\textit{[Collaboration demo]}
      \end{block}
  \end{columns}
\end{frame}