\section{PDDL Crate Navigation}
% ----- When A* Fails ---------------------------------------------
\begin{frame}{When A* Fails}
    \begin{itemize}
        \item A* fails: no path that avoids crates
        \item Second A* ignoring crates finds a ``ghost'' path
        \item Crates intersecting that path are part of the ``PDDL`` problem instance
    \end{itemize}
\end{frame}

% ----- PDDL  -------------------------------------------
\begin{frame}{PDDL}
  \begin{columns}
    \column{0.55\textwidth}
      \begin{itemize}
        \item Cluster detected $\to$ PDDL problem built from current beliefs
        \item Solver finds \texttt{move} + \texttt{push} sequence to reach goal
        \item Result translated back to agent's action plan
      \end{itemize}
    \column{0.43\textwidth}
      %#TODO: uncomment this for video demo
      %\animategraphics[autoplay,loop,width=\linewidth]{12}{../../media/PDDL/frame-}{1}{256}
      \begin{block}{}
        \centering\textit{[PDDL demo]}
      \end{block}
  \end{columns}
\end{frame}

% ----- The PDDL Domain -----------------------------------------
\begin{frame}{The PDDL Domain}
  \begin{itemize}
    \item STRIPS: two object types - \texttt{tile} and \texttt{crate}
    \item 4 directional \textbf{move} actions (step into crate-free tile)
    \item 4 directional \textbf{push} actions:
    \begin{itemize}
      \item Agent adjacent on opposite side of push direction
      \item Destination must be a valid, crate-free tile
    \end{itemize}
    \item Push relocates crate; agent steps onto its former tile
  \end{itemize}
\end{frame}

% ----- Performance Bottlenecks ----------------------------------
\begin{frame}{Performance Bottlenecks}
  \begin{itemize}
    \item \textbf{Real-Time Constraints:} The initial PDDL solver integration was too slow to meet the game's strict timing demands.
    \item \textbf{Identified Issues \& Fixes:}
    \begin{itemize}
      \item \emph{API Overhead:} Remote solver calls introduced heavy latency. \textbf{Fix:} Migrated to a local solver instance.
      \item \emph{Redundant Computation:} Identical crate configurations were being solved repeatedly, but implementing a standard cache was non-trivial.
    \end{itemize}
    \item \textbf{The Pivot:} Rather than brute-forcing a complex caching mechanism, we restated the problem to naturally maximize cache hits.
  \end{itemize}
\end{frame}

% ----- Smart Fix -----------------------------------------------
\begin{frame}{Optimization: Problem Restatement}
  \begin{block}{Three-Phase Journey Segmentation}
    \begin{enumerate}
      \item \textbf{Approach:} A* pathfinding from the current position to the cluster \emph{entry}.
      \item \textbf{Manipulation:} PDDL execution solely through the crate cluster (entry $\to$ exit).
      \item \textbf{Completion:} A* pathfinding from the cluster \emph{exit} to the final goal.
    \end{enumerate}
  \end{block}
\end{frame}

% ----- Smart Fix (Result) --------------------------------------
\begin{frame}{Optimization: Problem Restatement}
  \begin{columns}
    \column{0.4\textwidth}
      \begin{itemize}
        \item \textbf{The Result:} The PDDL problem is now strictly anchored to the cluster entry tile rather than absolute map coordinates.
        \item \textbf{Simplified Cache Key:} $\langle \text{entry},\, \text{exit},\, \{\text{crate positions}\} \rangle$.
      \end{itemize}
    \column{0.55\textwidth}
        \centering
        \resizebox{\linewidth}{!}{\input{../../figures/pddl-segments.tex}}
  \end{columns}
\end{frame}
